Cognate beginnings to bilingual lexical acquisition

### 2023-07-25 manuscript/manuscript.qmd
Fix Makefile and change Docker folder

The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in adults. For  instance, unbalanced (or low-proficiency) bilingual adults benefit more strongly from forward (dominance to non-dominant) cross-language phonological priming during word processing than from backward phonological priming (non-dominant to dominant) [e.g., @grainger1998masked; @voga2007cognate; @de1991lexical; @dimitropoulou2011masked]. On the contrary, more balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked]. Similarly, the infants in the present study benefited from the cognate status of words in their less dominant language, but not in their more dominant language. It is possible that in both cases, the underlying cause behind the asymmetric effect of parallel activation lies in lexical representations in the non-dominant language receiving stronger activation from their translations in the dominant language than vice versa.

This dominance-mediated difference in connection strength is contemplated by previous models of bilingual lexical access, such as the Revised Hierarchical Model [RHM; @kroll1994category], but the specific implementation of the source of such difference is still unclear.  A more recent proposal, the Ontogenic Model of L2 Lexical Representation [@bordag2022ontogenesis; @cook2016fuzzy], points to a 'fuzzier'---or more poorly phonological specification---of lexical representations in the non-dominant second language, compared to those in the dominant language. This sub-optimal specification may be due to a mismatch between the phonological inventory of both languages [@cutler2006asymmetric; @sebastian2005influence; @sebastian2006first] or, more generally, an imprecise encoding of the features of the representation in any domain (e.g., form-meaning mapping, orthographical, phonological). Although this account seems convenient to explain cross-language asymmetries in language processing, it should be taken into account that most (if not all) of the proposals aforementioned are aimed at explaining results in adults, and not in children. Language dominance and proficiency are frequently defined in both populations using different dimensions. For instance, the low proficiency bilinguals in many of the aforementioned studies acquired their second language years after their toddlerhood. This might have changed substantially the developmental trajectory of language acquisition they followed, compared to the bilinguals in the present study, who were exposed to both languages from early. Moreover, our findings can in principle be explained in the absence of a poorer specification of representations in the non-dominant language. As suggested in the introduction, it is possible that simply the fact that infants are exposed more frequently to words in the dominant language leads to words in the non-dominant language receiving additional parallel activation, compared to words in the dominant language. This explanation does not necessarily compete with words in the dominant language activating more strongly their translations than vice versa, and further research is needed in order to clarify this issue.

>>>>

The asymmetric facilitation of cognateness on word acquisition reported in the present study parallels previous findings in adults. For  instance, unbalanced (or low-proficiency) bilingual adults benefit more strongly from forward (dominant to non-dominant) cross-language phonological priming during word processing than from backward (non-dominant to dominant) phonological priming [e.g.; @grainger1998masked; @voga2007cognate; @de1991lexical; @dimitropoulou2011masked]. On the contrary, more balanced (or high-proficiency bilinguals) show an equivalent priming facilitation in both directions [@basnight2007differences; @dunabeitia2009masked]. Similarly, the infants in the present study benefited from the cognate status of words in their less dominant language, but not in their more dominant language. It is possible that, in both cases, that such an asymmetric effect of parallel activation is the result of lexical representations in the non-dominant language receiving stronger activation from their translations in the dominant language than vice versa. This dominance-mediated difference in connection strength is contemplated by previous models of bilingual lexical access, but these models differ in the specific implementation of the source of the asymmetry. Some models hard-code such asymmetries, so that the connections between word representation sand other words or their concepts in one language are stronger than those of word representations in the other language [e.g., the Revised Hierarchical Model; @kroll1994category]. In other models, like the Multilink model [@dijkstra2019multilink], lexical representations in the non-dominant language have a lower resting activation due to the lower frequency at which their word-forms are used. Some other models implement a combination of the former. For instance, the BIA/BIA+ [@dijkstra2013bia; @dijkstra2002architecture] models implemented a combination of (1) lower frequency---in a similar fashion than Multilink---, together with (2) a stronger top-down inhibition of word representations in the non-dominant language.

More recent proposals capture some developmental features of the bilingual lexicon, and suggest that structural and function asymmetries between the dominant and the non-dominant language may arise from simply a difference in the frequency of the input, the age of acquisition of the non-dominant language. For instance, simulations of subsequent bilinguals from the DevLex-II model [@zhao2010bilingual; @zhao2013simulating] result in an asymmetric cross-language priming, in which words from the dominant (earlier acquired) language primed more strongly the recognition of words in the non-dominant language (later acquired) than in the other direction. In addition, the simulated lexica showed sparser connectivity between words in the non-dominant language, compared to the connectivity between words in the dominant language.

A more recent proposal, the Ontogenic Model of L2 Lexical Representation [@bordag2022ontogenesis; @cook2016fuzzy], points to a 'fuzzier'---or more poorly phonological specification---of lexical representations in the non-dominant second language, compared to those in the dominant language. This sub-optimal specification may be due to a mismatch between the phonological inventory of both languages [@cutler2006asymmetric; @sebastian2005influence; @sebastian2006first] or, more generally, an imprecise encoding of the features of the representation in any domain (e.g., form-meaning mapping, orthographical, phonological). Although this account seems convenient to explain cross-language asymmetries in language processing, it should be taken into account that most (if not all) of the proposals aforementioned are aimed at explaining results in adults, and not in children. Language dominance and proficiency are frequently defined in both populations using different dimensions. For instance, the low proficiency bilinguals in many of the aforementioned studies acquired their second language years after their toddlerhood. This might have changed substantially the developmental trajectory of language acquisition they followed, compared to the bilinguals in the present study, who were exposed to both languages from early. Moreover, our findings can in principle be explained in the absence of a poorer specification of representations in the non-dominant language. As suggested in the introduction, it is possible that simply the fact that infants are exposed more frequently to words in the dominant language leads to words in the non-dominant language receiving additional parallel activation, compared to words in the dominant language. This explanation does not necessarily compete with words in the dominant language activating more strongly their translations than vice versa, and further research is needed in order to clarify this issue.
